---
title: "Week 6"
author: "Shraddha Somani"
date: "6/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Transformation

```{r nycflights13}
library(nycflights13)
```

```{r tidyverse}
library(tidyverse)
```

To do data manipulation we are using flights data

## Filter

```{r flights}
flights
# select all flights for January 1
(jan1 <- filter(flights, month == 1, day == 1))
# select all flights for november december
(nov_dec <- filter(flights, month %in% c(11, 12))) 
# De-Morgan's Law
filter(flights, !(arr_delay > 120 | dep_delay > 120))
filter(flights, arr_delay <= 120, dep_delay <= 120)
```

## Arange Rows
arrange works similarly to filter instead of selecting rows, it changes their order

```{r}
arrange(flights, year, month, day)
arrange(flights, desc(dep_delay))
```

## Select
It helps in narrowing in on the variables we’re actually interested in

```{r}
# Select columns by name
select(flights, year, month, day)
# Select all columns from year to day
select(flights, year:day)
# Select all columns minus from year to day
select(flights, -(year:day))
# Rename a variable
rename(flights, tail_num = tailnum)
# Everything is useful if we have few variables and we would like to move to the start of data frame
select(flights, time_hour, air_time, everything())
```

## Mutate
Adding new variable with mutate. Columns are always added in the end of dataset

```{r}
# Select data set
flights_sml <- select(flights, year:day, ends_with("delay"), distance, air_time )
# Mutate
mutate(flights_sml, gain = dep_delay - arr_delay, speed = distance / air_time * 60)
mutate(flights_sml, gain = dep_delay - arr_delay, hours = air_time / 60, gain_per_hour = gain / hours)
# If we only want new variables
transmute(flights, gain = dep_delay - arr_delay, hours = air_time / 60, gain_per_hour = gain / hours)
# transmutate with modular function
transmute(flights, dep_time, hour = dep_time %/% 100, minute = dep_time %% 100)
```

## Summaries
```{r}
# summarise with group by is highly useful
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
# Combining multiple operations with pipe
delays <- flights %>% group_by(dest) %>% summarise( count = n(), dist = mean(distance, na.rm = TRUE), delay = mean(arr_delay, na.rm = TRUE)) %>%filter(count > 20, dest != "HNL")
# Missing Values
not_cancelled <- flights %>% filter(!is.na(dep_delay), !is.na(arr_delay))
not_cancelled %>% group_by(year, month, day) %>% summarise(mean = mean(dep_delay))
# Planes with highest average delays
delays <- not_cancelled %>% group_by(tailnum) %>% summarise(delay = mean(arr_delay))
ggplot(data = delays, mapping = aes(x = delay)) + geom_freqpoly(binwidth = 10)
# Useful summary functions like mean meadian min max
not_cancelled %>% group_by(year, month, day) %>% summarise(avg_delay1 = mean(arr_delay),avg_delay2 = mean(arr_delay[arr_delay > 0]))
# Standard Deviation
not_cancelled %>% group_by(dest) %>% summarise(distance_sd = sd(distance)) %>% arrange(desc(distance_sd))
# Min Max to measure Rank
not_cancelled %>% group_by(year, month, day) %>% summarise(first = min(dep_time),last = max(dep_time))
# destinations with most carriers
not_cancelled %>% group_by(dest) %>% summarise(carriers = n_distinct(carrier)) %>%  arrange(desc(carriers))
# Count for destination
not_cancelled %>% count(dest)
# Grouping of multiple variables
daily <- group_by(flights, year, month, day)
(per_day   <- summarise(daily, flights = n()))
(per_month <- summarise(per_day, flights = sum(flights)))
# Ungrouping
daily %>% ungroup() %>% summarise(flights = n())
```

## Grouped Mutate
We vcan use mutate with filter

```{r}
# Worst member of each group
flights_sml %>% group_by(year, month, day) %>% filter(rank(desc(arr_delay)) < 10)
# Groups bigger than threshold
(popular_dests <- flights %>% group_by(dest) %>% filter(n() > 365))
# Standardised per group metrics
popular_dests %>% filter(arr_delay > 0) %>% mutate(prop_delay = arr_delay / sum(arr_delay)) %>% select(year:day, dest, arr_delay, prop_delay)
```


