---
title: "Week7"
author: "Shraddha Somani"
date: "6/18/2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Visualization

```{r tidyverse}
library(tidyverse)
```

```{r ggplopt2}
library(ggplot2)
```

To do data visualization we are using mpg data

## Creating ggplot

```{r mpg}
mpg
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))
```

```{r mapdata}
library(mapdata)
```

## Aesthetic mapping
We can do changes in color shape size of variables while plotting

```{r}
# Color to point different class of car
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = class))
# Size to map class will give warning because of mapping an unordered variable to ordered variable
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, size = class))
# Alpha aesthetioc for transparency
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
# Shape aesthetic to control shape of the points
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, shape = class))
# We can also set aesthetic manually like here color is set to red
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), color = "red")
```

## Common Problem
common problem when creating ggplot is + in wrong place

```{r}
## ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))
```

## Facets
Another way to add variable is to split your plot into facets, subplots that each display one subset of the data.

```{r}
# Facet_wrap to facet plot by a single variable
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_wrap(~ class, nrow = 2)
# Facet plot on combination of 2 variables
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(drv ~ cyl)
```

## Geometric Objects
We can change geometric objects as well there are over 30 geom available

```{r}
# Point Geom
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))
# Smooth Geom: a smooth line fitted to the data
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy))
# We can even change the line type
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
# We can change legend as well which will give multiple smooth line
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, color = drv), show.legend = FALSE)
# We can also add multiple geom to same plot
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + geom_smooth(mapping = aes(x = displ, y = hwy))
# We can use same approach for class data changing aesthetics and everything
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + geom_point(mapping = aes(color = class)) + geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```

## Statistical Transformation


```{r diamonds}
# Bar chart for diamond dataset grouped on cut
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut))
# Algorithim used is statistical so we can use stat count as well
ggplot(data = diamonds) + stat_count(mapping = aes(x = cut))
# map height of the bars to the raw values of y variable 
#stat of geom bar is changed from default to identity
demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)
ggplot(data = demo) + geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")
# Overide the default mapping with transformed variables
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
# stat summary to draw attention to statistical tranformation
ggplot(data = diamonds) + stat_summary(mapping = aes(x = cut, y = depth),fun.ymin = min, fun.ymax = max,fun.y = median)
```

## Position adjustment
We can change color of a bar char or do other aethetic changes

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = cut))
# Multiple variable stacking
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = clarity))
# Position change of stacking to see overlapping
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + geom_bar(alpha = 1/5, position = "identity")
ggplot(data = diamonds, mapping = aes(x = cut, colour = clarity)) + geom_bar(fill = NA, position = "identity")
# Dodge places overlapping object beside each other
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```

## Coordinate system
We can change the coordinates as well although default is cartesian

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + geom_boxplot()
# coordinate change
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + geom_boxplot() + coord_flip()
# Coord quick map helps setting aspect ratio perfectly
nz <- map_data("nz")
ggplot(nz, aes(long, lat, group = group)) + geom_polygon(fill = "white", colour = "black")
ggplot(nz, aes(long, lat, group = group)) + geom_polygon(fill = "white", colour = "black") + coord_quickmap()
# coord polar is used for polar coordinateswhich reveals connection between bar chart and coxcomb chart
bar <- ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = cut), show.legend = FALSE, width = 1) + theme(aspect.ratio = 1) + labs(x = NULL, y = NULL)
bar + coord_flip()
bar + coord_polar()
```



