---
title: "Week10"
author: "Shraddha Somani"
date: "7/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Clustering

```{r packages}
library(tidyverse)
library(cluster)
library(factoextra)
library(gridExtra)
library(dendextend)
```

To do data manipulation we are using USArrests Data

```{r USArrests}
df <- USArrests
# Remove missing value
df <- na.omit(df)
# standardize/scale data because we dont want clustering algo to depend on arbitrary value
df <- scale(df)
head(df)
```

## K Means clustering
Center means number of cluster and nstart meansnumber of configuration

```{r}
k2 <- kmeans(df, centers = 2, nstart = 25)
str(k2)
k2
# visualize cluster
fviz_cluster(k2, data = df)
# If more than 2 dimensions fviz_cluster will do PCA
# we can also do scatter plot
df %>% as_tibble() %>% mutate(cluster = k2$cluster, state = row.names(USArrests)) %>% ggplot(aes(UrbanPop, Murder, color = factor(cluster), label = state)) + geom_text()
```

## Cluster comparison
we can change centers and compare the cluster

```{r}
k3 <- kmeans(df, centers = 3, nstart = 25)
k4 <- kmeans(df, centers = 4, nstart = 25)
k5 <- kmeans(df, centers = 5, nstart = 25)
# plots to compare
p1 <- fviz_cluster(k2, geom = "point", data = df) + ggtitle("k = 2")
p2 <- fviz_cluster(k3, geom = "point",  data = df) + ggtitle("k = 3")
p3 <- fviz_cluster(k4, geom = "point",  data = df) + ggtitle("k = 4")
p4 <- fviz_cluster(k5, geom = "point",  data = df) + ggtitle("k = 5")
# Lets arrange multiple plots side by side
grid.arrange(p1, p2, p3, p4, nrow = 2)
```

## Optimal Cluster
There are different ways to find optimal number of cluster

```{r}
set.seed(123)
# elbow method
# since bent appears to be on 4 optimal number of cluster is 4
fviz_nbclust(df, kmeans, method = "wss")
# Average silhoutee method
# It shows 2 as cluster choice 
fviz_nbclust(df, kmeans, method = "silhouette")
# Gap Statistics Method shows 4 as optimal cluster size
gap_stat <- clusGap(df, FUN = kmeans, nstart = 25, K.max = 10, B = 50)
fviz_gap_stat(gap_stat)
```

## Compute result with 4 cluster
```{r}
set.seed(123)
final <- kmeans(df, 4, nstart = 25)
print(final)
fviz_cluster(final, data = df)
```

## Hierarchical Clustering
There are 2 type AGNES and DIANA

# Agglomerative clustering (AGNES)
```{r}
# Hclust function
# Dissimilarity matrix
d <- dist(df, method = "euclidean")
# Hierarchical clustering using Complete Linkage
hc1 <- hclust(d, method = "complete" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
# Agnes function in this we get agglomerative coefficient which measure amount of clustering structure
hc2 <- agnes(df, method = "complete")
# Agglomerative coefficient
hc2$ac
# This helps us to find stronger clustering method
# check different methods
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")
# compute coefficient
ac <- function(x) {agnes(df, method = x)$ac}
map_dbl(m, ac)
# Plot agnes dendogram
hc3 <- agnes(df, method = "ward")
pltree(hc3, cex = 0.6, hang = -1, main = "Dendrogram of agnes") 
```

# Diana Clustering

```{r}
# Diana clustering
hc4 <- diana(df)
# Divise coefficient
hc4$dc
# plot dendrogram
pltree(hc4, cex = 0.6, hang = -1, main = "Dendrogram of diana")
```